Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-30 paper/ms.tex
new abstract, new conclusions

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host
one LAE with a probability $f_{\rm occ}$.  We find that the observational information on the number density restricts the minimum mass to a
range $10^{10}\hMsun < M_{\rm min}< 10^{11.5}\hMsun$ leaving all the
other parameters unrestricted, while the clustering information puts a
tight constraint of $M_{\rm   max}<10^{12}\hMsun$. This gives support
to the idea that the most massive dark matter halos at that epoch do
not host the brightest LAEs.  However, none of these observations is
able to constrain the occupation fraction. It is always possible to
find a model that is consistent with observations and has any desired
occupation fraction in the range $0.1\leq f_{\rm occ}\leq
1.0$. We also find that all models have a narrow mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ smaller than 1.0 dex,
with the vast majority in the range $\Delta M<0.5$ dex, meaning that
there is only a factor of $3$ in mass between the minimum and maximum
halo mass. We divide the best models into families of low and high occupation
fraction with a transition around $f_{\rm occ}\leq 0.3$. These two
families are almost disjoint in their values of $M_{\rm min}$. Low
occupation fraction occurs below a threshold of $M_{\rm th}=
10^{10.8}\Msun$, while high occupation fractions are present in
models with $M_{\rm min}>M_{\rm th}$. This represents an unconclusive response as
to which is the occupation fraction for LAEs at $z=3.1$. A definite
answer requires a detailed physical model for the  connection between
dark matter, star formation and the escape of \ly radiation. That kind
of modelling should also provide a plausible explanation for the
narrow range in mass $\Delta M<1.0$ dex derived here for the halos
hosting LAEs at $z=3.1$.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. For each halo in the box we store its
position in the box (3-D coordinates) and FOF mass. We focus our work
on halos more massive than $1\times 10^{10}$\hMsun that are resolved
with at least $70$ particles, the reasons for this choice are
explained in the next sub-section.

@@

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying attention to the impact of cosmic variance in these results. We
build a large number of mock catalogs matching observational
geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos with a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE. We proceed with a
thorough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ 
combinations of parameters in the model we find $49$ arrangements
consistent with observations.

The successful parameter arrangements present a wide range of possibilities for
DM hosting LAE hosts at $z=3.1$. This means that the available LAEs' spatial
information is not sufficient to put a tight constraint on the mass
and occupation fraction of DM halos hosting LAEs. In particular it is
not possible to put a strong constraint on the occupation
fraction. Such a wide range in solutions is facilitated by the large
dispersion in the statistics derived from the mocks. All the halo mass
ranges and occupation fractions deduced in previous analysis in
previous analysis \citep[i.e.][]{Gawiser2007,Ouchi2010} are a subset of the
models we find in this \documentname.

Nevertheless, the successful models can be split into three families
depending the mass range $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm min}$ and the occupation fraction $f_{\rm occ}$. The
first family is narrow both in $\Delta M$ and $f_{\rm occ}$, a second
family is only narrow in $\Delta M$ and the third is narrow in $f_{\rm
  occ}$ and broad in $\Delta M$.

We find a large number of models with a narrow mass range
$\Delta M<1.0$ dex that are consistent in every aspect with the
spatial distribution of observed LAEs. In order to have this characteristic
it is required that halos around the mass $M_{\rm   max}$ for each
model become inefficient in hosting detectable LAEs. Two conditions
can explain this. The first is having LAEs with a decreasing \ly
escape fraction with increasing mass, the second is having larger screening
effects by neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

A third aspect that stands out is the existence of a dominant family
of models with narrow $\Delta M<1.0$, as described before, and also a
low occupation fraction $f_{\rm occ}<0.3$. This gives support to the
common wisdom stating that only a small fraction of star forming
galaxies can be detected as strong LAEs.

@@

In summary, we show how current observations of LAEs cannot
put a tight constraint on the physical models. We  foresee that the
new observations with new instruments (such as MUSE, Hyper SuprimeCam
and HETDEX) covering larger fields and a wider range of luminosities
will be key in imposing tighter constraints on the properties of dark
matter halos hosting LAEs.

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. The authors also thank Peter Laursen, Paulina Lira, 
Alvaro Orsi and Mark Dijkstra for helpful comments on the physical
interpretation and presentation of our results.

>>>>

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the geometries of observed fields. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host
one LAE with a probability $f_{\rm occ}$. This main motivation to
build such a model is to avoid the large uncertainties in the estimation of
\ly luminosities. We find that the clustering and number density information are
insufficient to impose a tight constraint on the escape fraction. On
the other hand the minium mass and maximum mass are tightly
constrained to be $M_{\rm max}<10^{12}$\hMsun and
$10^{10}\hMsun\leq M_{\rm min}\leq 10^{11.5}\hMsun$.  We also find
that all models have a narrow mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ smaller than 1.0 dex,
with the vast majority in the range $\Delta M<0.5$ dex. All previous
constraints on the mass range and occupation fraction are a sub-set of
the models we report in this paper. This represents an unconclusive response as
to which is the occupation fraction for LAEs at $z=3.1$. A more
precise answer requires physical models for the connection between
dark matter, star formation and the escape of \ly radiation. That kind
of modeling should also provide a plausible explanation for the narrow
 ange in mass $\Delta M<1.0$ dex derived here for the halos hosting
 LAEs at $z=3.1$.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{MultiDark}. For each halo in the box we store its
position in the box (3-D coordinates) and FOF mass. We focus our work
on halos more massive than $1\times 10^{10}$\hMsun that are resolved
with at least $70$ particles, the reasons for this choice are
explained in the next sub-section.

@@

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying special attention to the impact of cosmic variance on these
results. To this end we build a large number of mock catalogs matching
observational geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos with a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE. We proceed with a
thorough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ 
combinations of parameters in the model we find $23$ arrangements
consistent with observations.

We find that the observational information we use is insufficient to
impose a tight constraint on the escape fraction. On the other hand
the minium mass and maximum mass are tightly constrained as follows
$M_{\rm max}<10^{12}$\hMsun and $10^{10}\hMsun\leq M_{\rm min}\leq
10^{11.5}\hMsun$. Furthermore, we find that the mass range defined as
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ is always
$\Delta M\leq 1.0$.

The wide range in solutions is facilitated by the large dispersion in
the statistics derived from the mocks. Nevertheless, all the halo mass
ranges and occupation fractions deduced previous analysis
\citep[i.e.][]{Gawiser2007,Ouchi2010} are a subset of the models we
find in this \documentname. All of our models support the notion that
the most massive halos at $z=3.1$ do not host the brightest LAEs.

The existence of a narrow mass range for halos hosting LAEs, $\Delta
M\leq 1.0$ dex, has two different explanations at least. The
first is having LAEs with a decreasing \ly escape fraction with
increasing mass, the second is having larger screening effects by
neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

@@

In summary, we show how spatial information of LAEs cannot
put a tight constraint on their host halos and occupation fraction. We
foresee that the new observations with new instruments (such as MUSE,
Hyper SuprimeCam and HETDEX) covering larger fields and a wider range
of luminosities will be key in imposing tighter constraints on the
properties of dark matter halos hosting LAEs. At the same time,
additional modeling for \ly radiation transfer, is needed to put a
tighter constrain on the \ly escape fraction in high redshift
galaxies.

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. The authors also thank Peter Laursen, Paulina Lira, 
Alvaro Orsi and Mark Dijkstra for helpful comments on the physical
interpretation and presentation of our results. J.E.F-R was
supported by the FAPA grant by Vicerrector\'ia de Investigaciones at
Universidad de los Andes.

The MultiDark Database used in this paper and the web application
providing online access to it were constructed as part of the
activities of the German Astrophysical Virtual Observatory as result
of a collaboration between the Leibniz-Institute for Astrophysics
Potsdam (AIP) and the Spanish MultiDark Consolider Project
CSD2009-00064. The Bolshoi and MultiDark simulations were run on the
NASA's Pleiades supercomputer at the NASA Ames Research Center.
